---
title: "Basic Clustering On PheWAS Data"
author: "Nick Strayer"
date: "10/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
source("biSBM.R"); ### loading and compiling biSBM function
```

## Data

We're using a snapshot exported from the phewas explorer.

```{r}
network_data <- read_rds('data/testing_network_data.rds')
network_data$edges %>% head() %>% kable()
network_data$vertices %>% head() %>% kable()
```


## Running BiSBM on data

```{r, cached = TRUE}
node_clusters <- biSBM(
  data = network_data$edges, 
  nodeType = network_data$vertices$name %>% str_detect('case') %>% ifelse(1, 2), 
  ka = 5,       # Number of clusters desired for cases 
  kb = 3,       # Number of clusters desired for phenotypes
  deg.corr = 1, # Do we correct for node degrees in the algorithm? Yes due to smaller sizes of rarer-phenotypes
  iter = 5      # How many iterations? Algorithm picks the best one.
);

```


## Visualizing the results


### Network

First we can view this as a network diagram colored by cluster membership.

```{r}
colors <- c('#8dd3c7','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9','#bc80bd','#ccebc5')
new_node_colors <- purrr::map_chr(node_clusters, ~colors[.])

list(
  edges = network_data$edges,
  vertices = network_data$vertices %>% mutate(color = new_node_colors)
) %>% 
  jsonlite::toJSON() %>%
  r2d3::r2d3(
    script = 'd3_plots/network_2d.js',
    container = 'div',
    dependencies = "d3-jetpack",
    options = list(
      just_snp = FALSE, 
      msg_loc = 'message'
    )
  )
```


### Phenotype Profiles of Case Clusters

Next, we can look at the phenotype profiles by cluster membership. 

We first need to make a helper dataframe containing the new clusters. 
```{r}
clustered_vertices <- network_data$vertices %>% 
  mutate(
    type = ifelse(str_detect(name,'case'), 'case', 'phenotype'),
    cluster = as.integer(node_clusters)
  ) %>% 
  select(index, name, type, cluster, snp_status, tooltip) 
```

Now we can use this to construct a bar chart for each cluster of their phenotype entries. 

```{r}
phenotype_vertices <- network_data$vertices %>% filter(!str_detect(name,'case'))

get_phenotype_vec <- function(index){
  network_data$edges %>% 
    filter(source == index) %>% 
    mutate(target = as.integer(target)) %>% 
    left_join(phenotype_vertices, by = c('target' = 'id')) %>% 
    select(source, phenotype = name)
}

phenotypes_by_cluster <- clustered_vertices %>% 
  filter(type == 'case') %>%
  .$index %>%    
  purrr::map_df(get_phenotype_vec) %>% 
  mutate(source = as.integer(source)) %>% 
  inner_join(clustered_vertices %>% select(index, cluster), by = c('source' = 'index'))
  

phenotypes_by_cluster %>% 
  ggplot(aes(x = phenotype)) +
  geom_bar() +
  facet_grid(cluster~.) +
  ggtitle('Phenotype Profiles of Case Clusters')
```



